### 2023-01-05 supplementary_materials.Rmd

---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(data.table)
library(flextable)
library(magrittr)
library(table1)
```

# Supplementary Material

## NACC Exclusion Criteria

National Alzheimer's Coordinating Center (NACC) participants were excluded if they had any of the following conditions noted in the NACC Neuropathology Data Set. Variable names and descriptions are taken from <https://files.alz.washington.edu/documentation/rdd-np.pdf>. Variable descriptions may be lightly edited. Participants were not excluding for missing data in any of these fields.

```{r}
table_s1 <- fread("doc/nacc_exclusion_variables.txt", sep = "|") %>% 
  flextable() %>% 
  add_header_lines("Table S1: NACC Exclusion Criteria")

table_s1
```

\newpage

```{r}

pord <- fread("data/mega/mega_np_ord.pheno", na.strings = '-1')
pbin <- fread("data/mega/mega_np.pheno", na.strings = '-1')
pheno <- merge(pord, pbin, c("FID", "IID"))
covar <- fread("data/mega/mega_np.covar")
covar_apoe <- fread("data/mega/mega_np_apoe.covar", stringsAsFactors = TRUE)
covar <- merge(covar, 
               covar_apoe[, .(FID, IID, apoe)], 
               c("FID", "IID"), 
               all.x = TRUE)
covar <- covar[, .(FID, IID, msex, age_death, apoe)]
covar[, Sex := factor(msex, labels = c("Female", "Male"))]

dt <- merge(pheno, covar, by = c("FID", "IID"))
setnames(dt, "age_death", "Age of Death")
dt[, Study := factor(FID, labels = c("NACC", "ROSMAP", "ACT", "ADNI"))]
dt[, CERAD := ordered(cerad, 
                        labels = c("None", "Mild", "Moderate", "Severe"))]

# Display apoe variable as number of APOE e4 alleles for table
dt[apoe %in% c("apoe22", "apoe23", "apoe33"), 
     `APOE e4 alleles` := 0]
dt[apoe %in% c("apoe24", "apoe34"), 
     `APOE e4 alleles` := 1]
dt[apoe %in% c("apoe44"), 
     `APOE e4 alleles` := 2]
dt[, `APOE e4 alleles` := ordered(`APOE e4 alleles`)]

# add all other NPE used as outcomes
dt[, `Braak Stage` := ordered(braak)]
dt[, `Diffuse Amyloid` := 
     ordered(diffuse_abeta, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, Arteriolosclerosis := 
     ordered(arteriol, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, Atherosclerosis := 
     ordered(atheroscler, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, CAA := 
     ordered(caa_ord, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, `LATE-NC` := 
     ordered(late, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, `Lewy Body` := 
     ordered(lewy_body, 
             labels = c("None", "Mild", "Moderate", "Severe"))]
dt[, HS := factor(hs, labels = c("Absent", "Present"))]
dt[, Microinfarcts := factor(microinf, labels = c("Absent", "Present"))]
dt[, `Gross infarcts` := factor(grossinf, labels = c("Absent", "Present"))]

t1dt <- 
  as.data.frame(
    table1(~ Sex + `Age of Death` + `APOE e4 alleles` + CERAD + 
             `Braak Stage`+ `Diffuse Amyloid` + Arteriolosclerosis +
             Atherosclerosis + CAA + `LATE-NC` + `Lewy Body` + HS +
             Microinfarcts + `Gross infarcts`| Study,
           data = dt
           )
    )

t1 <- flextable(t1dt) %>% 
  autofit() %>% 
  padding(padding = 2, part = "all") %>%
  add_header_lines("Table S2: Participant Demographics") %>% 
  add_footer_lines(paste0("Key: SD, standard deviation; Min, minimum; ",
                          "Max, maximum. LATE-NC, LATE neuropathologic change",
                          " HS, Hippocampal sclerosis",
                          "Age of Death variable is integer ",
                          "in NACC, ADNI, and ACT but continuous in ROSMAP."))

# saveRDS(t1, 
#         file = "output/manuscript/tableS1_participant_demographics.Rds")
t1
```

\newpage

![Figure S1: Dendrogram of Neuropathology Endophenotype Polychoric Correlations](doc/figure_S1_dendrogram.png)